$Z^\prime$ mass limits and the naturalness of supersymmetry

### 2015-01-27 paper/e6ssm-tuning-paper.tex
More small fixes.

While this approach has many merits here we will employ the traditional
measure of fine tuning because it is both simple to apply and easy to
compare with previous results due to it's widespread use.  Fortunately
the derivatives which appear in these tuning are also similar to the
Bayesian motivated measure so there should not be too large a
discrepancy between the two approaches.

@@

\section{\label{sec:model}U(1) extensions and the E$_6$SSM}
In this paper we consider U(1) extensions of the MSSM where the gauge
group at low energies is, \be SU(3)_C\times SU(2)_W\times U(1)_Y\times
U(1)^\prime, \ee where $U(1)^\prime$ is the new gauge group beyond
that of the SM and MSSM. The minimal superfield content of $U(1)$
extensions which solve the $\mu$-problem should be ordinary left
handed quark, $\hat{Q}_i$ and lepton $\hat{L}_i$
($i=1,2,3$)superfields along with right handed superfields
$\hat{u}^c_i$, $\hat{d}^c_i$, $\hat{e}^c_i$ ($i=1,2,3$) for the
up-type (s)quarks, down-type (s)quarks and charged (s)leptons
respectively and three Higgs superfields, up-type $\hat{H}_u$,
down-type $\hat{H}_d$ and a singlet under the SM gauge group
$\hat{S}$.

@@

Here we will refer to U(1) extensions of the MSSM, which solve the
$\mu$-problem as the USSM \cite{Cvetic:1995rj, Jain:1995cb,
  Nir:1995bu, Cvetic:1996mf, Cvetic:1997ky}.  The couplings for the
$U(1)^\prime$ gauge group should allow the following renormalisable
superpotential terms required in the USSM, \be W_{USSM} = y^U_{ij}
\hat{u}^c_i \hat{H}_u \hat{Q}_j + y^D_{ij} \hat{d}^c_i \hat{H}_d
\hat{Q}_j + y^E_{ij} \hat{e}^c_i \hat{H}_d \hat{L}_j + \lambda \hat{S}
\hat{H}_d \hat{H}_u, \ee with $i,j \in \{1,2,3\}$.

@@

Here we will focus most on the special case where the gauge symmetry
is $U()_N$,under which the right handed neutrino remains
interaction-less. This is the case in the the exceptional
supersymmetric standard models (E$_6$SSM)
\cite{King:2005jy,King:2005my,Athron:2010zz}, and closely-related
variants\cite{Howl:2007zi, Braam:2009fi, Braam:2010sy, Hall:2011zq,
  Nevzorov:2012hs, Athron:2014pua}.  Since the right handed neutrino
has no gauge symmetry protecting it's mass from becoming extremely
heavy such models may explain the tiny observed masses of neutrinos
via the see-saw mechanism and the baryon asymmetry in the universe via
leptogenesis \cite{Hambye:2000bn,King:2008qb, King:2008gw}.  Recently
it has also been studied in the context of electroweak baryogenesis
\cite{Chao:2014hya}.

The gauge coupling running in the E$_6$SSM at the two-loop level leads
to unification more precisely than in the MSSM \cite{King:2007uj} or
in slightly modified scenarios two-step unification can take place
\cite{Howl:2007hq,Howl:2007zi}. If the exotic particles are light in
these models this can open up non-standard decays of the SM--like
Higgs boson \cite{Hall:2010ix,Nevzorov:2013tta,Athron:2014pua}. The
correct relic abundance could be obtained entirely through an almost
decoupled ``inert'' neutralino sector \cite{Hall:2009aj}.  However
this is no longer phenomenologically viable due to limits from direct
detection of dark matter
\cite{2011PhRvL.107m1302A,2012PhRvL.109r1301A, Akerib:2013tjd} and due
to a significant suppression of the decay of the lightest Higgs boson
into SM states, due to a new channel into inert singlinos opening up.

There are still several remaining options. One may specialize to
scenarios known as the EZSSM \cite{Hall:2011zq} where the inert
singlinos, which cause these problems are entirely decoupled and the
relic abundance is fitted with a bino-like candidate with a novel
mechanism involving back-scattering into a Heavier inert
Higgsino. Another well motivated scenario admits two possible dark
matter candidates \cite{Nevzorov:2012hs}, where one will be an inert
singlino and the other will have a similar composition to MSSM
neutralinos. The simplest phenomenologically viable solution in that
case is to make the singlinos extremely light hot dark matter
candidate and explains almost all the relic abundance density with

@@

Nonetheless while this model is very elegant so far, the
superpotential, (\ref{Eq:e6-superpot}) contains dangerous terms which
can induce proton decay and lead to large flavour changing neutral
currents (FCNCs).  There are a number of approaches to suppress these
terms, involving the use of different discrete symmetries.  Here for
the purposes of renormalisation group running we will simply include
the following unsuppressed superpotential terms, which follows the
approach taking in work on the cE$_6$SSM \cite{Athron:2009ue,
  Athron:2009bs},
%
%% \begin{align}
%%   \begin{split}
    \ba
    W &\approx& - y_{\tau} (\SuperField{H}_d
    \SuperField{L}_3) \SuperField{e}^C_3 - y_b (\SuperField{H}_d
    \SuperField{Q}_3) \SuperField{d}_3^C - y_t (\SuperField{Q}_3
    \SuperField{H}_u) \SuperField{u}_3^C\\
    &&
    + \lambda_i \SuperField{S} (\SuperField{H}_{1i}
    \SuperField{H}_{2i}) + \kappa_i \SuperField{S} \SuperField{D}_i
    \SuperField{\overline{D}}_i + \mu' (\SuperField{H}'
    \SuperField{\overline{H'}}),
    \ea
  %% \end{split}
  \label{SuPot_RGE}
%%\end{align}
%

@@

As stated above, in order to quantify the resulting fine tuning we

@@

of the fact that the amount of running involved is small. This makes
it possible to use approximate analytic solutions to the RGEs that
exhibit good accuracy over the range of scales considered. Given the
two-loop RG equation for a parameter $q$,

>>>>

While this approach has many merits here we will employ the traditional
measure of fine tuning because it is both simple to apply and easy to
compare with previous results due to it's widespread use.  Fortunately
the derivatives which appear in these tunings are also similar to the
Bayesian motivated measure so there should not be too large a
discrepancy between the two approaches.

@@

\section{\label{sec:model}U(1) extensions and the E$_6$SSM}
In this paper we consider U(1) extensions of the MSSM where the gauge
group at low energies is, \be SU(3)_C\times SU(2)_W\times U(1)_Y\times
U(1)^\prime. \ee $U(1)^\prime$ is the new gauge group beyond
that of the SM and MSSM. The minimal superfield content of $U(1)$
extensions which solve the $\mu$-problem should be ordinary left
handed quark, $\hat{Q}_i$ and lepton $\hat{L}_i$
($i=1,2,3$) superfields along with right handed superfields
$\hat{u}^c_i$, $\hat{d}^c_i$, $\hat{e}^c_i$ ($i=1,2,3$) for the
up-type (s)quarks, down-type (s)quarks and charged (s)leptons
respectively and three Higgs superfields, up-type $\hat{H}_u$,
down-type $\hat{H}_d$ and a singlet under the SM gauge group
$\hat{S}$.

@@

Here we will refer to U(1) extensions of the MSSM, which solve the
$\mu$-problem, as the USSM \cite{Cvetic:1995rj, Jain:1995cb,
  Nir:1995bu, Cvetic:1996mf, Cvetic:1997ky}.  The couplings for the
$U(1)^\prime$ gauge group should allow the following renormalisable
superpotential terms required in the USSM, \be W_{USSM} = y^U_{ij}
\hat{u}^c_i \hat{H}_u \hat{Q}_j + y^D_{ij} \hat{d}^c_i \hat{H}_d
\hat{Q}_j + y^E_{ij} \hat{e}^c_i \hat{H}_d \hat{L}_j + \lambda \hat{S}
\hat{H}_d \hat{H}_u, \ee with $i,j \in \{1,2,3\}$.

@@

Here we will focus most on the special case where the gauge symmetry
is $U(1)_N$, under which the right handed neutrino remains
interaction-less. This is the case in the the exceptional
supersymmetric standard models (E$_6$SSM)
\cite{King:2005jy,King:2005my,Athron:2010zz}, and closely-related
variants\cite{Howl:2007zi, Braam:2009fi, Braam:2010sy, Hall:2011zq,
  Nevzorov:2012hs, Athron:2014pua}.  Since the right handed neutrino
has no gauge symmetry protecting it's mass from becoming extremely
heavy such models may explain the tiny observed masses of neutrinos
via the see-saw mechanism and the baryon asymmetry in the universe via
leptogenesis \cite{Hambye:2000bn,King:2008qb, King:2008gw}.  Recently
it has also been studied in the context of electroweak baryogenesis
\cite{Chao:2014hya}.

The gauge coupling running in the E$_6$SSM at the two-loop level leads
to unification more precisely than in the MSSM \cite{King:2007uj} or,
in slightly modified scenarios, two-step unification can take place
\cite{Howl:2007hq,Howl:2007zi}. If the exotic particles are light in
these models this can open up non-standard decays of the SM--like
Higgs boson \cite{Hall:2010ix,Nevzorov:2013tta,Athron:2014pua}. The
correct relic abundance could be obtained entirely through an almost
decoupled ``inert'' neutralino sector \cite{Hall:2009aj}.  However
this is no longer phenomenologically viable due to limits from direct
detection of dark matter
\cite{2011PhRvL.107m1302A,2012PhRvL.109r1301A, Akerib:2013tjd} and due
to a significant suppression of the decay of the lightest Higgs boson
into SM states, due to a new channel into inert singlinos opening up.

There are still several remaining options. One may specialize to
scenarios known as the EZSSM \cite{Hall:2011zq} where the inert
singlinos, which cause these problems are entirely decoupled and the
relic abundance is fitted with a bino-like candidate with a novel
mechanism involving back-scattering into a heavier inert
Higgsino. Another well motivated scenario admits two possible dark
matter candidates \cite{Nevzorov:2012hs}, where one will be an inert
singlino and the other will have a similar composition to MSSM
neutralinos. The simplest phenomenologically viable solution in that
case is to make the singlinos extremely light hot dark matter
candidate and explains almost all the relic abundance density with

@@

Nonetheless while this model is very elegant so far, the
superpotential, (\ref{Eq:e6-superpot}) contains dangerous terms which
can induce proton decay and lead to large flavour changing neutral
currents (FCNCs).  There are a number of approaches to suppress these
terms, involving the use of different discrete symmetries.  Here for
the purposes of renormalisation group running we will simply include
the following unsuppressed superpotential terms, which follows the
approach taken in work on the cE$_6$SSM \cite{Athron:2009ue,
  Athron:2009bs},
%
%% \begin{align}
%%   \begin{split}
    \ba
    W &\approx& - y_{\tau} (\SuperField{H}_d
    \SuperField{L}_3) \SuperField{e}^C_3 - y_b (\SuperField{H}_d
    \SuperField{Q}_3) \SuperField{d}_3^C - y_t (\SuperField{Q}_3
    \SuperField{H}_u) \SuperField{u}_3^C\\
    &&
    + \lambda_i \SuperField{S} (\SuperField{H}_{1i}
    \SuperField{H}_{2i}) + \kappa_i \SuperField{S} \SuperField{D}_i
    \SuperField{\overline{D}}_i + \mu' (\SuperField{H}'
    \SuperField{\overline{H'}}),
    \ea
  %% \end{split}
  \label{SuPot_RGE}
%%\end{align}
%

@@

As stated above, to quantify the resulting fine tuning we

@@

of the fact that the running is over much smaller scales than when
evolving up to the GUT scale. This makes it possible to use
approximate analytic solutions to the RGEs that exhibit good accuracy
over the range of scales considered. Given the two-loop RG equation
for a parameter $q$,
